---
title: "Reproducible Research: Peer Assessment 1"
author: "UK"
date: "2016-6-16"
output: md_document
---

#Load the data
```{r}

library(knitr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(Hmisc)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)

if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
aData <- read.csv('activity.csv')
```

#What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
steps.date <- aggregate(steps ~ date, data = aData, FUN = sum)
```


2. Make a histogram of the total number of steps taken each day
```{r}
steps.date <- aggregate(steps ~ date, data = aData, FUN = sum)
barplot(steps.date$steps, names.arg = steps.date$date, xlab = "date", ylab = "steps")

```

3. Calculate and report the mean and median total number of steps taken per day
```{r}
mean(steps.date$steps)
median(steps.date$steps)
```


#What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=aData$steps), by=list(interval=aData$interval), FUN=mean, na.rm=TRUE)
```

1. Make a time series plot
```{r}
steps.interval <- aggregate(steps ~ interval, data = aData, FUN = mean)
plot(steps.interval, type = "l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```



# Imputing missing values
1. Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(aData))
```


2. Devise a strategy for filling in all of the missing values in the dataset.
```{r}
"I will use the means for the 5-minute intervals as fillers for missing values"
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newData <- merge(aData, steps.interval, by = "interval", suffixes = c("", ".y"))
nas <- is.na(newData$steps)
newData$steps[nas] <- newData$steps.y[nas]
newData <- newData[, c(1:3)]
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
steps.date <- aggregate(steps ~ date, data = newData, FUN = sum)
barplot(steps.date$steps, names.arg = steps.date$date, xlab = "date", ylab = "steps")

mean(steps.date$steps)

median(steps.date$steps)

```


# Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
activityDataImputed <- aData
activityDataImputed$steps <- impute(aData$steps, fun=mean)
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

2. Make a panel plot containing a time series plot
```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```


#End